---
title: "GP-Pipeline DB2TTL"
author: "Christian Ruiz"
date: "15 7 2021"
output: html_document
---

```{r setup, messages=F}
# ********************************
#Version 0.1 - 15.07.2021 - Raum, Zeit

# Variables to change
#ts: triple-store destination
ts<-"http://datashop.westeurope.cloudapp.azure.com/statistics/"


# Vorbereitungen ----
# *************************************
# Aendern der Sprache in R
Sys.setenv(LANG = "de")
# Clear R's Brain
#rm(list = ls())

# Laden der Libraries
library(pacman)
p_load(tidyverse, forcats, haven, readxl, stringr)

# Verbinden mit der DB ---> An persönliche EInstellungen anpassen
source("~/db_connect.R")
```


## Load GP Data

Load GP Data
# Insert here Code to load from DB

## Lets first start with Geo-Codes

We need first the property RAUM .../property/RAUM that we will try to conceive after the one by Statistik Stadt Zuerich (SSZ). 

Right now I do not understand why SSZ makes differences between e.g. https://ld.stadt-zuerich.ch/statistics/code/R00074 and codes like https://ld.stadt-zuerich.ch/statistics/code/ORT0663

But this includes also to create the schemes for Gemeinde, Bezirk, Planungsregion, Kanton, and Land (pointing as in SSZ also to Kontinent but not defining it...). Bezirk and Planungsregion are both children of Kanton (which has to be checked). 



```{r}
filecon<-file("data_ttl/raum_zeit.ttl",open="w",encoding="UTF-8")

writeLines(paste0("<",ts,"property/RAUM> a <http://www.w3.org/1999/02/22-rdf-syntax-ns#Property> ;\n","<http://schema.org/name> 'Raum' ;\n","<http://schema.org/identifier> 'RAUM' .\n\n"),filecon)

writeLines(paste0("<",ts,"scheme/Gemeinde> a <http://schema.org/DefinedTermSet> ;\n","<http://schema.org/name> 'Gemeinde' ;\n","<http://www.w3.org/2004/02/skos/core#broader> <",ts,"scheme/Bezirk> .\n\n"),filecon)

writeLines(paste0("<",ts,"scheme/Bezirk> a <http://schema.org/DefinedTermSet> ;\n","<http://schema.org/name> 'Bezirk' ;\n","<http://www.w3.org/2004/02/skos/core#broader> <",ts,"scheme/Kanton> .\n\n"),filecon)

writeLines(paste0("<",ts,"scheme/Planungsregion> a <http://schema.org/DefinedTermSet> ;\n","<http://schema.org/name> 'Planungsregion' ;\n","<http://www.w3.org/2004/02/skos/core#broader> <",ts,"scheme/Kanton> .\n\n"),filecon)

writeLines(paste0("<",ts,"scheme/Kanton> a <http://schema.org/DefinedTermSet> ;\n","<http://schema.org/name> 'Kanton' ;\n","<http://www.w3.org/2004/02/skos/core#broader> <",ts,"scheme/Land> .\n\n"),filecon)

writeLines(paste0("<",ts,"scheme/Land> a <http://schema.org/DefinedTermSet> ;\n","<http://schema.org/name> 'Land' ;\n","<http://www.w3.org/2004/02/skos/core#broader> <",ts,"scheme/Kontinent> .\n\n"),filecon)

   
   
```   

For that we will create a logic with codes that will use (also here in parallel to SSZ) an internal coding logic such as /code/R00132 and every node will contain extra information such as the name, the scheme of a regional element (Gemeinde, Bezirk and Planungsregion) and further. 

TODO: Differences to SSZ: We do not have yet a sameAs with wikidata such as <http://schema.org/sameAs> "http://www.wikidata.org/entity/Q392079" . And we also lack the position such as <http://schema.org/position> 10120000074 . It would also be an idea to have here a link to the BFS-NR. 

```{r}

for(i in 1:nrow(raum)){
  
  writeLines(paste0("<",ts,"code/R",str_pad(raum$RAUMEINHEIT_ID[i],5,pad="0"),"> a <http://www.w3.org/2004/02/skos/core#Concept> , <http://schema.org/Place> ;\n","<http://schema.org/name> '",raum$NAME[i],"'@de  ;\n","<http://www.w3.org/2004/02/skos/core#inScheme> <",ts,"scheme/",raum$DESCRIPTION[i],"> ;\n","<http://schema.org/identifier> 'R",str_pad(raum$RAUMEINHEIT_ID[i],5,pad="0"),"' .\n\n"),filecon)
}




```

## Time

And now we will add the time-elements. First of all we will define the properties ZEIT and TIME such as SSZ. Then there are somoe additional schemas hasEnd and referenceTime. We also need codes for certain temporal elements, but SSZ seem not to have defined them yet: Tag, Monat, Quartal, Trimester, Semester, Jahr.  

```{r}
writeLines(paste0("<",ts,"property/ZEIT> a <http://www.w3.org/1999/02/22-rdf-syntax-ns#Property> ;\n","<http://schema.org/name> 'Zeit' ;\n","<http://schema.org/description> 'Repräsentation der Zeit als Konzept, repräsentiert die vollständige Komplexität.' ;\n","<http://schema.org/identifier> 'ZEIT' .\n\n"),filecon)

writeLines(paste0("<",ts,"property/TIME> a <http://www.w3.org/1999/02/22-rdf-syntax-ns#Property> ;\n","<http://schema.org/name> 'Zeit' ;\n","<http://schema.org/description> 'Repräsentation der Zeit als ISO 8601 Datum (vereinfacht).' ;\n","<http://schema.org/identifier> 'TIME' .\n\n"),filecon)

writeLines(paste0("<",ts,"schema/hasEnd> a <http://www.w3.org/1999/02/22-rdf-syntax-ns#Property> ;\n","<http://schema.org/name> 'Ende der Periode'@de ;\n","<http://schema.org/description> 'Ende einer zeitlichen Einheit'@de ;\n","<http://schema.org/identifier> 'hasEnd' .\n\n"),filecon)

writeLines(paste0("<",ts,"schema/referenceTime> a <http://www.w3.org/1999/02/22-rdf-syntax-ns#Property> ;\n","<http://schema.org/name> 'Referenzzeit' .\n\n"),filecon)




```

We will define the min and max year of any indicator in the GP. Then, we will create for every year a ZEIT element such as SSZ. 


```{r}
for(i in min(indikator$MIN_JAHR,na.rm=T):max(indikator$MAX_JAHR,na.rm=T)){
 
  writeLines(paste0("<",ts,"code/Z3112",i,"> a <http://www.w3.org/2004/02/skos/core#Concept> , <http://www.w3.org/2006/time#TemporalEntity> ;\n",
  "<http://schema.org/name> 'Z3112",i,"' ;\n",
  "<",ts,"schema/hasEnd> '",i,"-12-31'^^<http://www.w3.org/2001/XMLSchema#date> ;\n",
  "<",ts,"schema/referenceTime> <",ts,"code/Tag> ,\n",
  "<",ts,"code/Monat> ,\n",
  "<",ts,"code/Quartal> ,\n",
  "<",ts,"code/Trimester> ,\n",
  "<",ts,"code/Semester> ,\n",
  "<",ts,"code/Jahr> .\n\n"),filecon)
  
}
close(filecon)
```

## Specifying the datasets

Instead of a string like in SSZ, we will use the ID of the Indikator with a DS in front for dataset. 

TODO: Here we have a lot of variables that we can add to the SSZ ones.
Do the cube.link elements have to be imported into the triple store?


```{r}
filecon<-file("data_ttl/datasets.ttl",open="w",encoding="UTF-8")
for(i in 1:nrow(indikator)){
  writeLines(paste0("<",ts,"DS",str_pad(indikator$INDIKATOR_ID[i],4,pad="0"),"> a <https://cube.link/Cube> ;\n",
                    "<http://schema.org/name> '",indikator$NAME[i],"' ;\n",
                    "<http://schema.org/identifier> 'DS",str_pad(indikator$INDIKATOR_ID[i],4,pad="0"),"' ;\n",
             "<https://cube.link/observationSet> <",ts,"DS",str_pad(indikator$INDIKATOR_ID[i],4,pad="0"),"/observation/> ;\n",
             "<https://cube.link/observationConstraint> <",ts,"DS",str_pad(indikator$INDIKATOR_ID[i],4,pad="0"),"/shape/> .\n\n"),filecon)
  
}

close(filecon)
```

## Looping over the observations

TODO: 
- MEASURES MUESSEN DEFINIERT WERDEN!!!
- WEITERE PROEPRTIES SUCH AS SEX OR HERKUNFT HAVE TO BE DEFINED
- Attribute Korrektur is missing
- As in SSZ we add XXX0000XXX0000XXX0000 to the variables - without knowing why 

```{r}

#code might be improved because of for loop

indikatorvalue$label<-paste0("DS",str_pad(indikatorvalue$INDIKATOR_ID,4,pad="0"),"-R",str_pad(indikatorvalue$RAUMEINHEIT_ID,5,pad="0"),"-Z3112",indikatorvalue$JAHR)



for(i in indikator$INDIKATOR_ID){
  filecon<-file(paste0("data_ttl/observationset_",str_pad(i,4,pad="0"),".ttl"),open="w",encoding="UTF-8")

  print(paste0("DATASET:",i))
  writeLines(paste0("<",ts,"DS",str_pad(i,4,pad="0"),"/observation/> a <https://cube.link/ObservationSet> ;\n<https://cube.link/observation> \n"),filecon)
  df<-indikatorvalue[indikatorvalue$INDIKATOR_ID==i,]
  
  
    #PART1: Observation Collection
   
    df$temp_string<-paste0("<",ts,"DS",str_pad(i,4,pad="0"),"/observation/",df$label,"> ,\n")
  
    #the last one has to be a point and not a comma
    df$temp_string[nrow(df)]<-str_replace(df$temp_string[nrow(df)],",\n",".\n\n")  
    
    writeLines(df$temp_string,filecon)
  
  close(filecon)
    filecon<-file(paste0("data_ttl/observations_",str_pad(i,4,pad="0"),".ttl"),open="w",encoding="UTF-8")
    #PART2: Observation Values
    
   writeLines(paste0("<",ts,"DS",str_pad(i,4,pad="0"),"/observation/",df$label,">  a <https://cube.link/Observation> ;\n",
                     "<",ts,"property/ZEIT> <",ts,"code/Z3112",df$JAHR,"> ;\n",
                     "<",ts,"property/RAUM> <",ts,"code/R",str_pad(df$RAUMEINHEIT_ID,5,pad="0"),"> ;\n",
                     "<",ts,"attribute/KORREKTUR> false ;\n",
                     "<",ts,"measure/M",str_pad(i,4,pad="0"),"> ",ifelse(is.na(df$VALUE),'"NA"',df$VALUE)," ;\n",
                     #TODO insert here properties
                     "<http://schema.org/identifier> 'Z3112",df$JAHR,"R",str_pad(df$RAUMEINHEIT_ID,5,pad="0"),str_pad(i,4,pad="0"),"XXX0000XXX0000XXX0000' ;\n",
                     "<https://cube.link/observedBy> <",ts,"> ;\n",
                     "<",ts,"property/TIME> '",df$JAHR,"-12-31'^^<http://www.w3.org/2001/XMLSchema#date> .\n\n" 
                    
                     ),filecon) 
    
    
  close(filecon)
  
  
  
  
}




```

Next steps: 
- Define the measures
- Define missing properties
- Use Subset, set and thema as a logical hierarchy
```{r}

```
