---
title: "Text to SQL - Questions generation for the Statbot-Challenge"
output: html_notebook
---
https://towardsdatascience.com/text-to-sql-learning-to-query-tables-with-natural-language-7d714e60a70d

```{r setup}
library(tidyverse)
```

# Generation of natural language questions & SQL Statements

1. Read in a dataset with indicator metadata and short descriptions which can be used to generate questions.

```{r}
indicators <-
  read_csv("input_data/EN_INDICATORS.csv") %>%
  left_join(read_csv("input_data/indicators_desc.csv"),
            by=c("INDICATOR_ID"="indicator_id"))
```
## Create a single dataset with all information

1. Create a combined dataset containing all the information (values, indicator labels, spatial unit labels)  to make the generation of the question / sql pairs easier 

```{r}
gp_data <- read_csv("input_data/EN_INDICATOR_VALUES.csv") %>% 
           left_join(indicators, by=c("INDICATOR_ID"="INDICATOR_ID")) %>% 
           rename(INDICATOR_NAME=NAME) %>% 
           left_join(read_csv("input_data/EN_T_SPATIALUNIT.csv"), by=c("SPATIALUNIT_ID"="SPATIALUNIT_ID")) %>%rename(UNIT_NAME=NAME) 
```
## Create a dataset with random values, years and municipalities per indicator

2. To fill in the questions and queries, random values, years and municipalities are drawn for each indicator. These are then integrated into the templates dynamically.

```{r}

spatial_units <- read_csv("input_data/EN_T_SPATIALUNIT.csv") %>% 
                  select(SPATIALUNIT_ID,rand_mun_name=NAME)

random_values <- gp_data %>% 
  filter(TYPE_ID==1) %>% 
  group_by(INDICATOR_ID, INDICATOR_NAME, short_desc, questiontype) %>% 
  summarize(rand_val=sample(VALUE,1), 
            rand_year=sample(YEAR,1),
            rand_mun=sample(SPATIALUNIT_ID,1)) %>% 
                        filter(questiontype<=2) %>% 
  left_join(spatial_units, by=c("rand_mun"="SPATIALUNIT_ID"))

## create vectors that are integrated into the templates

indicator <- random_values$short_desc

indicator_name <- random_values$INDICATOR_NAME

random_value <- random_values$rand_val

indicator_id <- random_values$INDICATOR_ID

indicator_year <- random_values$rand_year

municipality_id <- random_values$rand_mun

municipality <- random_values$rand_mun_name

# Use the vectors in glue templates

questions_and_queries <- tibble(
                          questions = c("how high is the {indicator} in {municipality}?", 
                                        
                                       "which municipality has the highest {indicator}?",
                                       
                                       "Which municipality has the minimum {indicator}?",
                                       
                                        "what are the highest, lowest and average {indicator}?",
                                       
                                        "How many municipalities have a {indicator} higher than {random_value} per year?",
                                       
                                        "How high is the total {indicator} in the Canton Zurich?",
                                       
                                        "Which Region had the lowest {indicator} in the year {indicator_year}?"
                                       ),
                         
                           queries=c("select iv.value from indicator_values2 iv
                                      join indicators i on iv.indicator_id = i.indicator_id
                                      join spatialunit s on iv.spatialunit_id = s.spatialunit_id
                                      where i.name like '{indicator_name}' and s.name like '{municipality}'", 
                                     
                                      "select s.name from indicator_values2 iv
                                      join indicators i on iv.indicator_id = i.indicator_id
                                      join spatialunit s on iv.spatialunit_id = s.spatialunit_id
                                      where i.name like '{indicator_name}'
                                     order by iv.value desc
limit 1",
                                     
                                      "select s.name from indicator_values2 iv
                                      join indicators i on iv.indicator_id = i.indicator_id
                                      join spatialunit s on iv.spatialunit_id = s.spatialunit_id
                                      where i.name like '{indicator_name}' order by iv.value asc
limit 1",
                                     
                                      "select max(iv.value), min(iv.value), avg(iv.value) from indicator_values2 iv
                                      join indicators i on iv.indicator_id = i.indicator_id
                                      where i.name like '{indicator_name}'",
                                     
                                      "select iv.year, count(*) from
                                      indicator_values2 iv
                                      join indicators i on iv.indicator_id = i.indicator_id
                                      join spatialunit s on iv.spatialunit_id = s.spatialunit_id
                                      where iv.VALUE > {random_value} and s.type_id = 1 and i.name like '{indicator_name}' group by iv.year",
                                     
                                      "select iv.value from indicator_values2 iv
                                      join indicators i on iv.indicator_id = i.indicator_id
                                      join spatialunit s on iv.spatialunit_id = s.spatialunit_id
                                      where s.type_id=8 and i.name like '{indicator_name}'",
                                     
                                      "select min(iv.value), s.name from indicator_values2 iv
                                      join indicators i on iv.indicator_id = i.indicator_id
                                      join spatialunit s on iv.spatialunit_id = s.spatialunit_id
                                      where s.type_id=4 and iv.year={indicator_year} and i.name like '{indicator_name}'"
                                      )

)

# generate the questions

dataframe <- tibble(
              question=purrr::map(questions_and_queries$questions, 
                                   ~glue::glue(.x)) %>% unlist(),
              sql=purrr::map(questions_and_queries$queries, 
                                    ~glue::glue(.x))%>% unlist()
              )

# save questions and queries
write.csv(dataframe,"questions_queries.csv")

# save table
# write.csv(gp_data %>% head(10), "gp_data.csv")

```


question tokens
```{r}
how high is the 
which municipality has the highest 
what are the highest, lowest and average 


which are the runif() comunes with the highest 
What is the average , minimum , and maximum 
Show municipalities with a ,  , over x and more than , , 
Show names for all municipalities except for those having a x 
Find the weight of the youngest dog
```

# Test the SQL-Queries

```{r}
library(RSQLite)
library(readr)
```

## Create DB
```{r}
# read in csvs
indicators<- read_csv("input_data/EN_INDICATORS.csv") 

indicator_values2 <- read_csv("input_data/EN_INDICATOR_VALUES.csv") 

spatialunit <- read_csv("input_data/EN_T_SPATIALUNIT.csv")

#create DB
conn <- dbConnect(RSQLite::SQLite(), "statbot.db")

dbWriteTable(conn, "indicators", indicators)
dbWriteTable(conn, "indicator_values2", indicatorvalues2)
dbWriteTable(conn, "spatialunit", spatialunit)

RSQLite::dbListTables(conn)
```

## Test the queries
```{r}
# read in csv containing the queries
queries <- read_csv("questions_queries.csv") 

# run all the queries
purrr::map(queries$sql, ~dbGetQuery(conn, .x))

dbGetQuery(conn, queries$sql[33])

```

```{sql connection=conn}
select s.name from indicator_values2 iv
join indicators i on iv.indicator_id = i.indicator_id
join spatialunit s on iv.spatialunit_id = s.spatialunit_id
where i.name like 'PW new registrations per 1000 inhabitants [amount]' 
order by iv.value asc
limit 1
```



